---
title: "Lab 5"
author: "Jordan Faires"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Lab 5 Objectives

- Learn to parse dates with "lubridate" package
- Get counts of observations with count()
- Uncount() rows 
- One and two-sample t-test
- Create a heatmap with geom_tile()

```{r, include = FALSE}
# Load packages & data 

library(tidyverse)
library(janitor)
library(here)

lobster_abundance <- read_csv(here::here("data", "lobster_abundance.csv"), 
                              na = "-99999") %>% 
  janitor::clean_names()

```

Use tidyr::uncount() function to convert our lobstre data from frequency format to case format

```{r}
lobster_tidy <- lobster_abundance %>% 
  tidyr::uncount(lobster_count)

```

#### Exploratory data visualization

We're only going to consider 'site' as our variable of interest 

```{r}

ggplot(lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site),
              alpha = .5,
              width = .2)

# Histogram where sites are separate
ggplot(lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free")

#Looks normal, but going to use quanitle quantile to see if it actually is 

ggplot(lobster_tidy, aes(sample = size_mm)) +
  geom_qq() +
  facet_wrap(~site)

```

#### Convert the 'date' column to class 'Date' 

We'll use 'lubridate' package to convert to date format and then to help us easily parse month and year

```{r}

lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date)
  )

```

Parse year and month using lubridate::month() & lubridate::year()

```{r}

lobster_parse_date <- lobster_date %>%
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE),
    obs_year = lubridate::year(date_new)
  )

```


Now lets find counts of observed lobsters based on different hierarchical groupings: 

First: let's count lobsters by year and month:

```{r}
lobster_ym <- lobster_parse_date %>% 
  dplyr::count(obs_year, obs_month)

lobster_y <- lobster_parse_date %>% 
  dplyr::count(obs_year)

lobster_site <- lobster_parse_date %>% 
  dplyr::count(site)

```

If we want to contain a summary table that contains statistics other than counts by group, it's easier to use group_by() + n()

```{r}

lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE),
    lobster_number = n()
  )

```

#### Find confidence intervals

Use t.test() function to find confidence intervals (for one sample) and perform t-tests to compare means of two samples

```{r}
ivee_lobster <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

t.test(ivee_lobster)

```


#### Two-sample t-test to compare means

We want to ask if there's a significant difference in lobster length at Naples & Mohawk reefs? 

We've done our necessary exploratory analyses to determine that a 2-sample for means comparison is appropriate. 

```{r}

napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mn_ttest <- t.test(mohk_sample, napl_sample)

```

There is a sigificant difference in lobster length between Naples and Mohawk Reef. 

Now we can do it with set of single vector.
```{r}

lobster_mn <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK"))

mm_test2 <- t.test(size_mm ~ site, data = lobster_mn)

```

In-line text references to statistics
```{r}
mm_test2$p.value
mm_test2$statistic

```

Here is the p-value: `r mm_test2$p.value`

Here is my t-statistic: `r mm_test2$statistic`

#### Now: a heatmap! 
```{r}

lobster_ys <- lobster_parse_date %>% 
  count(obs_year, site)

ggplot(data = lobster_ys, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))

```

